Journey: A Study of Scalabilty of Memory Management in Linux

### 2017-05-05 paper.tex
Fix page 9

The \texttt{rand} workload, on the other hand, writes to pages that are 
not subpages of the same huge page. As a result, a huge page
is allocated with almost every write, and system memory is exhausted and begins
swapping very quickly (20,000 operations compared to 4 million with the \texttt{cont} benchmark). Memory made available by kswapd is quickly consumed by
further allocations. Hence, the memory used by the workload continues to stay
close to its peak value. As expected, the behavior is the same irrespective of
the size of the swap file.

@@

\begin{figure}[t]
    \centering
    \begin{tabular}{c||c|c}
        & \% of ops & Time spent \\ \hline \hline
        Latency $\le$ 50ms & 99.97\% & 7.39s \\
        Latency $>$ 50ms & 0.03\% & 18.9s \\
    \end{tabular}
    \label{fig:latency_table}
\end{figure}

@@

In the \texttt{rand} benchmark, as before, all
memory is exhausted, and swapping begins very soon. Many memory accesses 
thus take milliseconds to complete
because they require paging in or out, which is slow due to disk access.
Even so, the majority of benchmark run time comes from a minority of accesses.
Figure \ref{fig:swap_time_rand_cdf} shows that as much as 95\% of memory
accesses occur on the micro-second scale. Of the remaining 5\%, almost all
accesses complete in less than 20ms. Moreover, notice that the size of the
swap file makes very little difference in this workload because the disk has to
do random seeks regardless of swap file size.

The \texttt{cont} benchmark has more surprising results.
The system only begins to
swap after about 4M operations, as before. Unlike the \texttt{rand} benchmark, we see a
number of operations that take hundreds of milliseconds. Figure
\ref{fig:swap_time_cont_cdf} shows that these only form a fraction of a percent
of all accesses done after swapping starts. However, Figure
\ref{fig:swap_time_cdf} shows that these few accesses form a majority of run
time; they impact performance significantly. Figure \ref{fig:latency_table} summarizes
this trend.

@@

As expected, the \texttt{rand} benchmark with hugepages on incurs the highest
total overhead since it begins swapping very early. For this benchmark we
observe an overhead of roughly 0.0002 jiffies per page memory touched after
swapping begins. Also, kswapd runs for longer periods of time with longer
intervals of sleep in between. This is most likely because the benchmark has
some probability of retouching a page, so free memory is used more slowly, but
kswapd also has to do more work to find pages that are inactive.

In contrast, the \texttt{cont} benchmark with hugepages disabled uses
computational time at roughly 0.0572 jiffies per page of memory touched.
kswapd runs more frequently and for shorter periods of time because new pages
are touched very quickly, requiring faster reclamation, but because all pages
are only touched once, kswapd has to do very little work to choose pages to
reclaim.

Notice also that the \texttt{cont} benchmark with hugepages enabled spends
surprisingly little time in kswapd. We suspect that kswapd runs once and choose
a large number of pages to swap out. Subsequently, both the workload and kswapd
wait for I/O. This seems consistent with our previous results, in which some
operations take on the order of seconds to complete, but we have not yet
verified it.

@@

The swapping system in Linux has complicated interactions with workloads and
transparent hugepages. It is not obvious how to optimize a workload for better
swapping behavior. Moreover, we see that under memory pressure, kswapd tends to
run more often during page faults than in its own context.

@@

The emergence of new technology has in the past often required operating systems
to change. New non-volatile memory technologies are making it possible to have
terabytes of memory. Recent changes to memory management in Linux demonstrate a
lot of focus on improving the scalability of its data structures
\cite{huang2016evolutionary}.

@@

RadixVM tries to overcome performance issues in highly
concurrent workloads due to serialization of memory management operations on
kernel data structures \cite{radixvm}. This work demonstrates that many parts of
existing memory management schemes are not scalable to larger systems.
Similarly, in our work, we examine scalability limitations of memory
management in the Linux kernel.

>>>>

The \texttt{rand} workload, on the other hand, writes to pages that are likely not
adjacent and likely not subpages of a single huge page. As a result, a huge page
is allocated with almost every write, and system memory is exhausted and begins
swapping very quickly (20,000 operations compared to 4 million with the \texttt{cont} benchmark). Memory made available by kswapd is quickly consumed by
further allocations. Hence, the memory used by the workload continues to stay
close to its peak value. As expected, the behavior is the same irrespective of
the size of the swap file.

@@

\begin{figure}[t]
    \centering
    \begin{tabular}{c||c|c}
        & \% of ops & Time spent \\ \hline \hline
        Latency $\le$ 50ms & 99.97\% & 7.39s \\
        Latency $>$ 50ms & 0.03\% & 18.9s \\
    \end{tabular}
\end{figure}

@@

The overall behavior is as expected. In the \texttt{rand} benchmark, as before, all
memory is exhausted quickly, and the benchmark induces swapping soon after the
benchmark begins. Many memory accesses thus take milliseconds to complete
because they require paging in or out, which is slow due to disk access.
Even so, the majority of benchmark run time comes from a minority of accesses.
Figure \ref{fig:swap_time_rand_cdf} shows that as much as 95\% of memory
accesses occur on the micro-second scale. Of the remaining 5\%, almost all
accesses complete in less than 20ms. Moreover, notice that the size of the
swap file makes very little difference in this workload because the disk has to
do random seeks regardless of swap file size.

The \texttt{cont} benchmark has more surprising results.
The system only begins to
swap after about 4M operations, as before. Unlike the \texttt{rand} benchmark, we see a
number of operations that take hundreds of milliseconds. Figure
\ref{fig:swap_time_cont_cdf} shows that these only form a fraction of a percent
of all accesses done after swapping starts. However, Figure
\ref{fig:swap_time_cdf} shows that these few accesses form a majority of run
time; they impact performance significantly.

@@

In the \texttt{rand} benchmark with hugepages enabled, we observe an overhead
of roughly 0.0002 jiffies per page of memory touched after swapping begins.
Also, kswapd runs for longer periods of time with longer intervals of sleep in
between. This is most likely because the benchmark has some probability of
retouching a page, so free memory is used more slowly, but kswapd also has to
do more work to find pages that have not been recently used.

In the \texttt{cont} benchmark with hugepages disabled, kswapd uses
0.0572 jiffies per page of memory touched.
kswapd runs more frequently and for shorter periods of time because new pages
are touched very quickly, requiring faster reclamation. But because all pages
are only touched once, kswapd has to do very little work to choose pages to
reclaim.

The \texttt{cont} benchmark with hugepages enabled spends
surprisingly little time in kswapd. We suspect that kswapd runs once and choose
a large number of pages to swap out. Subsequently, both the workload and kswapd
wait for I/O. This seems consistent with our previous results, in which some
operations take hundreds of milliseconds to complete, but we have not yet
verified it.

@@

The swapping system in Linux has complicated interactions with workloads and
transparent hugepages. It is not obvious how to optimize a workload for better
swapping behavior. Moreover, we see that under memory pressure, kswapd tends to
run more often during page faults than in its own context, causing very poor
workload performance.

@@

New memory technologies are making it possible to have
terabytes of memory. Recent changes to memory management in Linux demonstrate
focus on improving the scalability of its data structures
\cite{huang2016evolutionary}.

@@

RadixVM tries to mitigate performance issues in highly
concurrent workloads due to serialization of memory management operations on
kernel data structures \cite{radixvm}. This work demonstrates that many parts of
existing memory management schemes are not scalable to larger systems.
